---
title: "Untitled"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---


## R setup

```{r include=FALSE}
rm(list = ls())

library(here)
library(readr)
library(stringr)
library(sf)
library(tidycensus)
library(tigris)
library(ggplot2)
library(viridis)

source(here('code', "_dir_packages.R"))

dir.svi <- here('data', 'Social Vulnerability Index')

yr <- 2022
```


## data 

```{r - csv}

svi <- readr::read_csv(file = paste0(dir.svi, '/SVI_2022_US.csv'), show_col_types = F)


svi_sub <- svi %>%
  select(1:E_HH, matches("(POV150|UNEMP|NOHSDP|AGE65|MINRTY)", ignore.case = TRUE)) %>%
  ## remove `Flag` columns
  select(!starts_with('F_', ignore.case = F)) %>%
  ## remove margins of error
  select(-starts_with(c("M_", "MP"))) %>%
  select(-any_of(c('LOCATION', 'E_HU'))) %>%
  ## turn -999 into NA
  mutate(across(where(is.numeric), ~ na_if(., -999))) %>%
  as.data.frame()

svi_sub_skim <- skimr::skim(svi_sub)
```



```{r - gdb}

library(sf)

gdb <- paste0(dir.svi, '/SVI_2022_US/SVI2022_US_tract.gdb')

# 1) List layers (feature classes) in the GDB
st_layers(gdb)
#> shows layer names, geometry types, CRS, feature counts

# 2) Read one layer by name
svi_sf <- st_read(dsn = gdb, layer = "SVI2022_US_tract")  # example layer name

# 3) filter 
svi_sf_sub <- svi_sf %>%
  select(1:E_HH, matches("(POV150|UNEMP|NOHSDP|AGE65|MINRTY)", ignore.case = TRUE)) %>%
  ## remove `Flag` columns
  select(!starts_with('F_', ignore.case = F)) %>%
  ## remove margins of error
  select(-starts_with(c("M_", "MP"))) %>%
  select(-any_of(c('LOCATION', 'E_HU'))) %>%
  ## turn -999 into NA
  mutate(across(where(is.numeric), ~ na_if(., -999))) %>%
  ## filter mainland 
  filter(!ST_ABBR %in% c("AK","HI","DC","PR")) %>%
  # ----- pooled quintiles for all EP_* variables -----
  mutate(
    across(
      starts_with("EP_"),
      ~ {
        brks <- quantile(.x, probs = seq(0, 1, by = 0.2), na.rm = TRUE, type = 7)
        brks <- unique(brks)                       # handle duplicate quantiles
        if (length(brks) < 2) {
          # all values identical or mostly NA: return all-NA factor with Q1–Q5 levels
          factor(NA_real_, levels = paste0("Q", 1:5))
        } else {
          labs <- paste0("Q", seq_len(length(brks) - 1))
          cut(.x, breaks = brks, include.lowest = TRUE, labels = labs)
        }
      },
      .names = "{.col}_q"
    )
  )

```


```{r - plot - line}

CNTY_i <- '36061'
# var <- 'EPL_POV150'

names(svi_sf_sub)
vs <- stringr::str_subset(names(svi_sf_sub), "^EP")
vs


p <- svi_sf_sub %>%
  filter(STCNTY %in% c(CNTY_i, '06075')) %>%
  mutate(EPL_POV150 = EPL_POV150*100) %>%
  ggplot(.) +
  geom_point(aes(x=EPL_POV150, y=EP_POV150) ) +
  geom_abline(slope = 1, color = 'red') + 
  facet_wrap(.~STCNTY) + 
  theme_minimal()
p
```




  Use `EP_POV150` (the actual percent) as your modeling covariate when you want to estimate how responses to greening vary with poverty. It’s on a meaningful, comparable scale (percentage points).

  Use `EPL_POV150` (percentile rank) when you want to rank/stratify places (e.g., target top 20% most vulnerable) or for easy-to-read maps. Don’t treat it as linear in regressions—percentiles depend on the reference distribution and compress extremes.

  For cross-city/county comparisons, prefer using the *actual percentage (EP_POV150) and create your own, pooled quantiles/standardization* across all places you’re comparing. Avoid relying on EPL_POV150 because SVI percentiles can be reference-dependent (state vs. national) and aren’t guaranteed comparable across jurisdictions.
  
  
```{r - save data}

dir.gd <- 'G:/Shared drives/invest-health/data/SVI_2022_US_processed/'

svi_sf_sub_out <- svi_sf_sub %>%
  ## remove `EPL_` columns
  select(!starts_with('EPL_', ignore.case = F)) %>%
  ## remove raw data columns
  select(-any_of(c("E_POV150", "E_UNEMP", "E_NOHSDP", "E_AGE65", "E_MINRTY" )))

# write (create/overwrite file)
st_write(svi_sf_sub_out, dsn = paste0(dir.gd, 'SVI_2022_US_tract.gpkg'), layer = "SVI_2022_US_tract", delete_dsn = FALSE, delete_layer = TRUE)
# st_write(svi_sf_sub_out, dsn = paste0(dir.gd, 'SVI_2022_US_tract.shp'), delete_dsn = TRUE)
```


```{r - plot - hist}
# pick which columns to histogram: all numeric, or those starting with "EP_"
num_cols <- svi_sf_sub_out %>%
  sf::st_drop_geometry() %>%
  select(where(is.numeric) & starts_with("EP_"))  %>%
  names()

vars <- num_cols

df_long <- svi_sf_sub_out %>%
  sf::st_drop_geometry() %>%
  select(all_of(vars)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

ggplot(df_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "grey70", color = "white") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  labs(x = NULL, y = "Count", title = "Histograms across variables") +
  theme_minimal(base_size = 10) +
  theme(panel.grid = element_blank())

  
```


```{r - plot - map}

for (var in vs){
  p <- svi_sf_sub %>%
    filter(STCNTY %in% c(CNTY_i)) %>%
    # --- 2) plot with ggplot (facets by model) ---
    ggplot(.) +
    geom_sf(aes(fill = .data[[var]]), color = NA) +
    # facet_wrap(~ model, ncol = 2) +
    # facet_grid(rows = vars(period), cols = vars(model)) +
    scale_fill_distiller(
      palette = "YlOrRd",
      na.value = "grey90", direction = 1, guide = 'colourbar', 
      name = var
    ) +
  
    theme_minimal(base_size = 11) +
    theme(
      panel.grid = element_blank(), 
      axis.text = element_blank(),
      # legend.position = c(0.85, 0.02), # inside bottom-right; tweak or set "right"
      legend.position = "bottom",
      legend.box = "horizontal",
      # legend.key.size = unit(10, "pt"),
      legend.title = element_text(lineheight = 0.9)
    )
  
  # --- 3) save ---
  ggsave(
    filename = file.path(dir.figures, paste("svi_map_2022", CNTY_i, var, ".png", sep='_')),
    plot = p, width = 7, height = 5, units = "in", dpi = 600
  )

}
```



```{r - get tract for all states, eval=FALSE, include=FALSE}

library(tidycensus)
library(dplyr)
library(purrr)



## save dataframe as csv
f <- paste0("data/aoi/", "tract_pop_2023.csv")
tract_pop <- readr::read_csv(f)

```



```{r - clipped tract, eval=FALSE, include=FALSE}

dir.aoi   <- 'D:/data/shp/US_census_Cartographic_Boundary_Shapefiles/cb_2020_us_ua20_500k/cb_2020_us_ua20_500k.shp'
dir.tract <- 'D:/data/shp/US_census_Cartographic_Boundary_Shapefiles/cb_2019_us_tract_500k/cb_2019_us_tract_500k.shp'
dir.bg    <- 'D:/data/shp/US_census_Cartographic_Boundary_Shapefiles/cb_2019_us_bg_500k/cb_2019_us_bg_500k.shp'

ua <- st_read(dir.aoi) 
tracts <- st_read(dir.tract)  # Census Tracts
# bg <- st_read(dir.bg)  # Census Tracts


tracts <- st_make_valid(tracts)
# Calculate original tract area
tracts$tract_area <- st_area(tracts)

# # Ensure same CRS
# ua <- st_transform(ua, st_crs(tracts))

cities <- c("San Francisco", 
            # "New York", "Chicago", 
            "San Antonio", "Los Angeles", 
            "Minneapolis",
            "Seattle", "Pittsburgh")


ua.i <- ua %>%
  filter(str_detect(NAME20, paste(cities, collapse = "|")))
ua.i <- st_make_valid(ua.i)



# # Spatial filter: keep tracts that intersect any urban area polygon
# tract.ua.i <- tracts[st_intersects(tracts, ua.i, sparse = FALSE), ]
# # bg.ua.i    <- bg[st_intersects(bg, ua.i, sparse = FALSE), ]
# 
# tract.ua.i <- st_make_valid(tract.ua.i)


```



```{r - aoi}

f <- "./data/aoi/tract_ua_2020_py.shp"
aoi <- st_read(f) %>%
  select(-any_of(c("NAME", 'STATE_NAME', "LSAD", "NAMELSADCO", "AWATER", "tract_area", "overlap_ra", 'overlap_ar')))


states <- unique(aoi$STATEFP)

pop_2020 <- get_decennial(
  geography = "tract",
  variables = "P1_001N",  # Total population
  year = 2020,
  # sumfile = "pl",
  state = states,
  geometry = F
)



aoi_sf <- aoi %>%
  rename('NAME20_UA' = 'NAME20') %>%
  # filter(str_detect(NAME20, 'San Francisco')) %>%
  left_join(., pop_2020, by = c("GEOID" = "GEOID"))



```




