---
title: "Untitled"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}

rm(list = ls())

# install.packages('FedData')
library(FedData) # pal_nlcd()

library(here)

library(sf)
library(raster)
library(dplyr)
library(tmap)

dir.here <- here()

dir.output <- paste0(dir.here, "/", "results/output/")
```


## read raster
```{r}
bou <- st_read(dsn = paste0(dir.here, "/data/cb_2019_us_county_500k_06075_clip.shp"))

pop  <- raster(x = paste0(dir.here, "/data/usa_ppp_2020_UNadj_constrained_SF_proj_NoDataTo0.tif"))
lulc <- raster(x = paste0(dir.here, "/data/nlcd_2019_land_cover_SF.tif"))

supply <- raster(paste0(dir.output, "urban_nature_supply_9m2_500m_masked.tif"))
demand <- raster(paste0(dir.output, "urban_nature_demand_9m2_500m.tif"))

balance_total  <-raster(paste0(dir.output, "urban_nature_balance_totalpop_9m2_500m.tif"))
balance_percap <-raster(paste0(dir.output, "urban_nature_balance_percapita_9m2_500m.tif"))


crs(bou)
crs(balance_percap)


library(rgdal)
bou_sp <- as_Spatial(bou)
# crs(bou_sp) <- crs(balance_percap)
bou_sp <- spTransform(bou_sp, CRS(proj4string(balance_percap)))

crs(bou_sp)
crs(balance_percap)

balance_percap_clip <- crop(balance_percap, bou_sp, mask = T)

## crop and mask
r2 <- crop(balance_percap, extent(bou_sp))
balance_percap_clip <- mask(balance_percap, bou_sp)
# plot(balance_percap_clip)

lulc <- mask(lulc, bou_sp)
```



## plot raster

*Discrete* gradient options are "cat", "fixed", "sd", "equal", "pretty", "quantile", "kmeans", "hclust", "bclust", "fisher", "jenks", "dpih", "headtails", and "log10_pretty". 

*Continuous* gradient options are "cont", "order", and "log10". The first maps the values of col to a smooth gradient, the second maps the order of values of col to a smooth gradient, and the third uses a logarithmic transformation. 


```{r}
p_pop <-
  tm_shape(bou) +
  tm_polygons(col = "gray") +
  tm_shape(pop) +
  tm_raster(alpha = .6, palette = "YlOrRd", style = "kmeans", title = "Count") +
  tm_compass(type = "arrow", position = c(.83, .85)) +
  tm_scale_bar(breaks = c(0, 1, 2), position = c(0.8, .75), text.size = .8) +
  tm_layout(
    main.title = "Population",
    legend.position = c("left", "top"),
    inner.margins = 0.1
  )
```


```{r}
# Load in the NLCD legend, colors and descriptions from package FedData. 


legend<-pal_nlcd()
# legend

# lc_lable <- c("Open Water", "Developed Open", "Developed Low Int.", "Developed Med. Int.", 
#               "Developed High Int.", "Barren", "Deciduous Forest", "Evergreen Forest", 
#               "Mixed Forest", "Pasture/Hay", "Cultivated Crop", "Woody Wetland", "Emergent Wetland")
# 
# lc_color <- c("deepskyblue", "brown1", "brown2", "brown3", "brown4", "darkgrey", "darkolivegreen", "darkolivegreen1", 
#                 "darkseagreen1", "goldenrod3", "darkorange", "darkorchid1", "darkorchid4")

lc_lable <- legend$Class
lc_color <- legend$Color
lc_color <- c("#5475A8", "#FFFFFF", "#E8D1D1", 'white', 'white', 'white', "#D2CDC0", "#85C77E", "#38814E", 
              "#D4E7B0", "#AF963C", "#DCCA8F", "#FDE9AA", "#D1D182", "#A3CC51", "#82BA9E", "#FBF65D", "#CA9146",
              "#C8E6F8", "#64B3D5")


p_lc <-
  tm_shape(bou) +
  tm_polygons(col = "white") +
  tm_shape(lulc) +
  tm_raster(
    alpha = .6, 
    style = "cat",
    labels = lc_lable,
    palette = lc_color, colorNA = NULL, 

    title = "LULC"
  ) +
  tm_compass(type = "arrow", position = c(.83, .85)) +
  tm_scale_bar(breaks = c(0, 1, 2), position = c(0.8, .75), text.size = .8) +
  tm_layout(
    main.title = "Nature area",
    legend.position = c("left", "top"),
    inner.margins = 0.1
  )
p_lc
```


```{r}
p_s <- 
  tm_shape(bou) +
  tm_polygons(col="gray")+
  tm_shape(supply) +
  tm_raster(alpha = .6, palette = "YlGn", style = "kmeans", title = "m2/capita") +
  tm_compass(type = "arrow", position = c(.83, .85)) +
  tm_scale_bar(breaks = c(0, 1, 2), position = c(0.8, .75), text.size = .8) +
  
  # tm_credits("Data from UNA", position = c(.87, .03)) +
  # tm_layout(title = "Supply", title.size = 1.5, title.position = c("left", "top")) +
  tm_layout(
    main.title = "Supply",
    legend.position = c("left", "top"),
    inner.margins = 0.1) 

p_d <- 
  tm_shape(bou) +
  tm_polygons(col="gray")+
  tm_shape(demand) +
  tm_raster(alpha = .6, palette = "YlOrBr", style = "kmeans", title = "m2") +
  tm_compass(type = "arrow", position = c(.83, .85)) +
  tm_scale_bar(breaks = c(0, 1, 2), position = c(0.8, .75), text.size = .8) +
  tm_layout(
    main.title = "Demand",
    legend.position = c("left", "top"),
    inner.margins = 0.1) 


p_balance <- 
  tm_shape(bou) +
  tm_polygons(col="gray")+
  tm_shape(balance_total) +
  tm_raster(alpha = .6, palette = "RdYlGn", midpoint = 0,
            style = 'order', legend.reverse = T,
            # style = "kmeans", 
            title = "m2") +
  tm_compass(type = "arrow", position = c(.83, .85)) +
  tm_scale_bar(breaks = c(0, 1, 2), position = c(0.8, .75), text.size = .8) +
  tm_layout(
    main.title = "Balance",
    legend.position = c("left", "top"),
    inner.margins = 0.1) 


p_balance_percap <- 
  tm_shape(bou) +
  tm_polygons(col="gray")+
  tm_shape(balance_percap_clip) +
  tm_raster(alpha = .6, palette = "RdYlGn", midpoint = 0, 
            style = 'order', legend.reverse = T,
            title = "m2/capita") +
  tm_compass(type = "arrow", position = c(.83, .85)) +
  tm_scale_bar(breaks = c(0, 1, 2), position = c(0.8, .75), text.size = .8) +
  tm_layout(
    main.title = "Balance per capita",
    legend.position = c("left", "top"), 
    inner.margins = 0.1) 


ps <- tmap_arrange(p_pop, p_lc, p_balance_percap, 
                   p_d,   p_s,  p_balance, nrow = 2)
tmap_save(ps, filename = paste0(dir.here, "/figures/UNA.jpeg"), height=9, width=16, units="in", dpi=300)
```
